---
title: "Project 3: Prediction of 2016 Election"
author: "Group 8: Jackie Lee, Juan Gomez, Patrick Lyons, Ross Miglin"
resource_files:
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
require(data.world)
require(MASS)
require(ISLR)
require(dplyr)
require(ggplot2)
require(hexbin)
require(jsonlite)
require(shiny)
require(rsconnect)
require(knitr)
require(lubridate)
require(rmarkdown)
require(tidyr)
library(rsconnect)
library(class)
knitr::opts_chunk$set(echo = TRUE)
```

## **R Session Info**
```{r}
sessionInfo()
```

## **Introduction**
In order to analyze the methods discussed in class for model comparison, we decided to use data on the 2016 Presidential Election in order to make predictions of which candidate one based on different races and different occupations. This is especially interesting because we now know what factors where key in the success Trump. So we can make models and assumptions based on that information to see if these factors are as significant in our models as the were in real life.


## **Requiring data and initializing session**
We have retrieved our data and required the proper libraries in order to correctly compute our analysis. Our data can be viewed at https://data.world/juandi/f-17-edaproject-3.
```{r}
require(data.world)

project <- "https://data.world/juandi/f-17-edaproject-3"

df <- data.world::query(
  data.world::qry_sql("SELECT * FROM electiondata"),
  dataset = project
)

library(class)
library(glmnet)
library(leaps)
```
## **Simple and Multiple Linear Regressions

We started our analysis by choosing one or more variables to plot vs. candidates and adding trendlines to those plots. We first focused on how one's occupation might affect their voting preference.

```{r}
#Setting categorical variables to numeric in order to graph output
df$candidate[df$candidate == "Trump"] <- 0
df$candidate[df$candidate == "Clinton"] <- 1
df$candidate = as.numeric(df$candidate)

plot(formula = candidate~management_occup, data=df)
abline(lm(formula=candidate~management_occup, data=df),col="red",lwd=1.5)
fitOccup = lm(forumla=candidate~management_occup, data=df)
fitOccup
summary(fitOccup)
```
Since the dependent variable of our data is binary (0 for Trump, 1 for Clinton), laying a linear regression line on the data doesn't highlight valuable correlation. We graphed the residuals against the fitted data to resolve this issue.
```{r}
#Residuals Plot
fitOccup2 = lm(management_occup~agri_occup+candidate,data=df)
fitOccup2
points(management_occup,fitted(fitOccup2),col="blue",pch=20)
summary(fitOccup2)
plot(fitOccup2)
```
Because our R-squared value was still low, we decided to factor in other occupations in addition to management.
```{r}
#Non-linear Plot
fitOccup3 = lm(management_occup~agri_occup*candidate,data=df)
fitOccup3
plot(fitOccup3)
points(management_occup,fitted(fitOccup3),col="blue",pch=20)
summary(fitOccup3)
plot(fitOccup3)
```
## **K-Nearest Neighbors**

One of the methods used to conduct the analysis is K-Nearest Neighbors. Below shows the code and results for the analysis on this method.

First of all, we must set the training data to conduct KNN. We use the votes variable since the amount of votes  per state has no real effect on who won.
```{r}
# Setting training data
train = df$votes > 1000000
```


Our first test using KNN is using the white_pop and unemployment variable to predict a winner. As seen in the election, the white and unemployed voters were one of the demographics that largely voted for Trump.
```{r}
## Test 1: Using the white population percentage and unemployment population percentage

#Binding predictors
test1 = cbind(df$white_pop,df$unemployment)

#Running KNN test
knn.pred1 = knn(test1[train,], test1[!train,], df$candidate[train], k=1)
```

```{r}
#Confusion matrix and mean
table(knn.pred1,df$candidate[!train])
mean(knn.pred1==df$candidate[!train])
```


```{r}
## Test 2: Using the white population percentage and unemployment population percentage
test2 = cbind(df$latino_pop,df$management_occup)

#Running KNN test
knn.pred1 = knn(test2[train,], test2[!train,], df$candidate[train], k=1)

#Confusion matrix and mean
table(knn.pred1,df$candidate[!train])
mean(knn.pred1==df$candidate[!train])
```

## **Linear Discriminant Analysis**
Now, we want to predict the candidate using LDA. In this model, we create a sample for our training set.

```{r}
train2 <- sample(1:50, 25)
candidates2=lda(candidate~graduates_degree+asian_american_pop+bachelors+pct_dem,data=df,prior=c(1,1)/2,subset=train2)
candidates2
renderPlot({plot(candidates2)})
lda.pred2=predict(candidates2, df[-train2, ])
dflda2 = data.frame(lda.pred2)
table(lda.pred2$class,df[-train2, ]$candidate)
mean(lda.pred2$class==df[-train2, ]$candidate)
```
## **Quadratic Discriminant Analysis**
We want to add a little more flexibility to our covariance matrix by extending our model to QDA:

```{r}
qda.fit = qda(candidate~management_occup+service_occup+sales_occup+agri_occup,data=df)
qda.fit

df.33 = subset(df,management_occup>=33)
qda.class = predict(qda.fit, df.33)
table(qda.class$class,df.33$candidate)
mean(qda.class$class==df.33$candidate)
```

**Interactive Histogram and Boxplot**
```{r}
sliderInput("bins", "Number of bins:", min = 10, max = 50, value = 30)


renderPlot({
  x <- dflda2$LD1
  bins <- seq(min(x), max(x), length.out = input$bins + 1)
  ggplot(dflda2) + geom_histogram(mapping = aes(x=LD1), breaks = bins, color="blue") + facet_wrap(~class)})

renderPlot(ggplot(dflda2) + geom_boxplot(mapping = aes(x=class, y=LD1), color = "blue"))
```

**Using Predictors from KNN Best Subset Regression**
```{r}
candidates3=lda(candidate~pct_rep+less_than_high_school+asian_american_pop+unemployment,data=df,prior=c(1,1)/2,subset=train2)
candidates3
lda.pred3=predict(candidates3, df[-train2, ])
dflda3 = data.frame(lda.pred3)
table(lda.pred3$class,df[-train2, ]$candidate)
mean(lda.pred3$class==df[-train2, ]$candidate)
```

**Conclusion**
First we ran every single predictor alone through LDA, and then picked the predictors that we believed to be the best at predicting candidate. From there, we ran the LDA model for the four predictors that we picked. Next we wanted to compare the results of our model to the predictors from the best subset regression that are in the KNN analysis. After running those four predictors through the LDA model, we came to the conclusion that the predictors we picked are slightly better at predicting the candidate because of the higher mean accuracy rate from our model. All in all, our LDA model strongly predicts the right candidate.
